IRS County-to-County Migration Data

### 2018-08-23 manuscript_file.Rmd
pushed with some edits

---
output: 
  pdf_document:
    number_sections: true
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: LATEX/svm-latex-ms.tex
title: "IRS County-to-County Migration Data"
thanks: "The data and code that supports the creation of this data are available in the Supplementary Materials and online at  https://github.com/mathewhauer/IRS-migration-data."
author:
  - name: Mathew E. Hauer ^1,2^*
    affiliation: Florida State University
  - name: James M. Byars ^3^
    affiliation: University of Georgia
# abstract: ""
keywords: "Migration, IRS, Data, R"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
#fontfamily: mathpazo
fontsize: 12pt
spacing: single
bibliography:  LATEX/mybibfile.bib
biblio-style: apsr
# use apsr or nature
header-includes:
- \usepackage[all]{nowidow}
- \usepackage{rotating}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead{}
- \fancyhead[CE,CO]{IRS Migration Data}

@@

**OBJECTIVE**: To produce a single, flat data file containing complete county-to-county IRS migration flow data.

@@

**CONTRIBUTION**: To encourage and facilitate the use of this data, we provide the full R script to download and flatten the IRS migration data as counts and a finalized data set covering the period 1990-2010.

@@

# Introduction
Migration flow data (ie, the number of migrants from location *i* to location *j*) is typically difficult to obtain [@willekens2016international; @rogers2010indirect]. In the United States, migration flow data is available from three primary sources, depending on time period: the Decennial Census, the American Community Survey, and the Internal Revenue Service's (IRS) county-to-county migration data^[The IRS migration data are available at https://www.irs.gov/statistics/soi-tax-stats-migration-data]. Our focus here is on the IRS migration data.

The IRS began using tax data to estimate migration in the 1970s and 1980s [@engels1981measuring;@franklin2006pandora]. Beginning with tax year 1991 (migration year 1990), the IRS produces these data in conjunction with the US Census Bureau using the IRS Individual Master File which contains every Form 1040, 1040A, and 1040EZ. A migrant is identified when a current year's tax return contains an address that is different from the matched preceding years' return. For the 2002 tax year, the IRS migration data contained approximately 130 million returns [@gross2005internal]. If an address is identical over two years, a filer is considered a non-migrant. The IRS provides data on the number of returns (approximating the number of households) and the number of tax exemptions (approximating the number of individuals).

The IRS annual series of county-to-county migration data cover 95 to 98 percent of the tax filing universe (or approximately 87% of US households [@molloy2011internal]) and their dependents making these data the largest migration data source for count flows between counties in the United States. But since the IRS derived migration information from tax returns, those who do not file taxes are most likely to be underrepresented [@gross2005internal; @dewaard2016population], namely undocumented populations, the poor, the elderly, and college students [@gross2005internal]. However, the overwhelming majority of householders file US tax returns in the United States [@molloy2011internal].

Because of the annual availability, relatively long time series (the series starts in 1990), large universe, and long history of use, the IRS data are an attractive data source for conducting migration research (e.g. [@curtis2015recovery;@molloy2011internal;@shumway2001spatial;@frey2009great]). Unfortunately, these data exist in seven legacy formats, split between 2,000+ files making analysis with this data rather burdensome and has prevented the widespread use of this valuable resource.

To encourage and facilitate the use of this tremendous migration resource, we flattened the IRS migration data into a single, standardized file containing all migration flows for the period 1990-2010. By publishing both the R code to collate, process, and flatten the IRS migration data as well as the migration data itself, our hope is to save time and effort for other migration scholars and facilitate their use. Scholars who wish to use these data should still familiarize themselves with the strengths and weaknesses, idiosyncrasies, and design of these data (see [@gross2005internal; @engels1981measuring;@franklin2006pandora;@pierce2015soi] for discussions on the IRS data)and with the procedures outlined in this document and in the corresponding R code^[While the IRS migration data presently continues to 2015, we limit the data to the period 1990-2010. Starting in tax year 2012 (migration year 2011), the IRS introduced a change in their processing [@pierce2015soi] creating a break in the time series].

We have attempted to introduce as little post-processing as possible to process the data into a common format and we outline the single mathematical calculation we performed on the data below. Briefly, the IRS suppresses any dyadic migration flow that contains fewer than 10 tax filers. We calculate the difference between the total number of migrants and the total number of enumerated migrants (those flows containing 10 or more tax filers) and create a new origin/destination code for these unenumerated migrants.

The following document is organized as follows. First, we describe the differences in the seven legacy formats and the creation of FIPS 99999 representing "unknown" destination/origin migrants. Then, we discuss the parts of the R code used to download the IRS migration data and finally process the data into a common format. We then conclude with some important usage notes. Importantly, the datafile we create represents EXEMPTIONS as opposed to RETURNS to better mimic the number of individuals migrating rather than the number of households.

# Data Preparation

The IRS migration data for the period 1990-2010 are available in seven legacy formats. **Table 1** summarizes some of the similarities and differences in these formats. For every year, the IRS publishes approximately 104 data files. (52 state entities by in/out-migration. Some years contain .csv and .dat summary files.) The underlying file organization, file format, naming schema, and coding can differ between these legacy formats. Migration years 1990 and 1991 are available as fixed-width text files, while 1992-2010 are available as excel files. For years 1990-2003, the IRS separated in/out migration into separate folders while 2004-2010 are published in a single folder. Each legacy format utilizes a different file naming scheme as well, making pattern matching of file names (called grepping) difficult. Importantly, the IRS treats non-migrants and total migrants differently in the seven legacy formats. For 1990 and 1991, the IRS simply has a field that reads "County Non-Migrants" for non-migrants; for 1992-1994, the IRS introduced a State code 63 but two difference County codes (010 for 1992 and 1994 and 050 for 1993) creating a 5-digits FIPS code of 63010 or 63050. After 1995, the IRS smartly set the origin FIPS equal to the destination FIPS for non-migrants. Lastly, Total Migrants are treated differently too. For 1990 and 1991, the destination field simply reads "Total Migration." For 1992-1994, the IRS introduced a State Code 00 and county code 001 for total migrants. After 1995, the IRS used State Code 96 and county Code 000 for a combined 5-digit FIPS code of 96000.

The differences described above and in **Table 1** are only some of the differences that are of interest to the data we produce here. Total Migrants, ie FIPS 96000 for migration data after 1995, is also broken down into Total Mig - US (FIPS 97000), Total Mig - US Same State (FIPS 97001), Total Mig - US Diff St (FIPS 97003), and Total Mig - Foreign (FIPS 98000). The IRS did not code these migration flows in this manner for all years, and in some cases (such as Total Mig - Foreign) migration flows are not reported. For simplicity and data continuity purposes, we simply create a new origin/destination (FIPS 99999) that contains all unspecified migration flows. We do this by subtracting the number of enumerated migrants from migration flows with greater than 10 migrants from the total number of migrants. This way, the sum of all enumerated migrants in our dataset will equal the total number of migrants in the IRS dataset.

@@

| Years      | Data Format | File Organization         | Sample File naming | Treatment of non-migrants                     | Treatment of Total Migrants        |
|------------|-------------|---------------------------|--------------------|-----------------------------------------------|------------------------------------|
| 1990-1991  | txt         | Separate in/out migration | C9091pao.txt       | Destination field reads 'County Non-Migrants' | Destination field reads 'Total Migration'                  |
| 1992, 1994 | xls         | Separate in/out migration | C9293Pao.xls       | State code = 63, County code = 010            | State code = 00, County code = 001 |
| 1993       | xls         | Separate in/out migration | co934pao.xls       | State code = 63, County code = 050            | State code = 00, County code = 001 |
| 1995-2003  | xls         | Separate in/out migration | co956paor.xls      | Origin FIPS = Destination FIPS                | State code = 96, County code = 000 |
| 2004-2006  | xls         | Single folder             | co0405PAo.xls      | Origin FIPS = Destination FIPS                | State code = 96, County code = 000 |
| 2007-2008  | xls         | Single folder             | co0708oPa.xls      | Origin FIPS = Destination FIPS                | State code = 96, County code = 000 |
| 2009-2010  | xls         | Single folder             | co0910oPA.xls      | Origin FIPS = Destination FIPS                | State code = 96, County code = 000 |
Table: Select differences in the file formats, file organizations, naming, and treatment of various migration statistics.

# R Code

@@

## Data Download
The script `001-download_data.R` will download and unzip the migration data from the IRS' websites into a folder standardized format into subdirectory `MigData/.` The IRS data is in two primary formats: 1990-2003 and 2004-onwards. There are eight files that the IRS includes in their zip archives that contain no data (these are in years 1998, 1999, 2000, and 2001). After being downloaded and unzipped, these files are then deleted. If they are not deleted, they will cause the subsequent `for loops` to fail in the next section. These files do not contain any migration information, their names suggest they represent aggregation of migration flows (for example 'co990usi.xls' suggests county 99-2000 US in-migration) and we unsure exactly why these files are included or their purpose.

## Data Processing
The third and final section contains several `foreach` parallel processing loops to process the seven legacy formats into a common data format. These files are then row-bound using `rbindlist` and transformed into a 'short' data frame. **Table 2** demonstrates the general file layout. In- and out-migration files are processed separately and only unique dyadic flows are kept in the file flat file.

# Usage Notes
Any origin-destination pair with fewer than 10 tax filers over the entire period is excluded from the final datafile since no data would be recorded in the IRS datafile. These unenumerated flows are collected in the FIPS code 99999.

The count data come from the `exemptions` field of the IRS migration data. The original IRS migration data contains two consistent fields across all years of data: a `returns` field and an `exemptions` field. Returns are the number of tax returns filed while exemptions are a proxy for the members of the household.

US Counties are fairly stable geographic units but some changes in counties do exist. To try and keep as close to the original data fidelity as possible, we did not recode any geographic changes. For instance, Broomfield County CO (FIPS 07014) was created out of parts of Adams, Boulder, Jefferson, and Weld counties and thus contains data only after 2002. Users should be aware of any changes in geography or FIPS codes that could substantially alter any analyses^[More detailed information about county boundary, name, or FIPS changes can be found at the following locations https://www.census.gov/geo/reference/county-changes.html
http://www.nber.org/asg/ASG_release/County_City/FIPS/FIPS_Changes.pdf
https://www.cdc.gov/nchs/data/nvss/bridged_race/County_Geography_Changes.pdf
https://www.ddorn.net/data/FIPS_County_Code_Changes.pdf].

Users should be aware that we prepared these data to stay as close to the original data fidelity as possible, documented their creation, and with open-source computer code. These data should be used only with full awareness of the inherent limitations of the IRS migration data and with the knowledge of the procedures outlined in this document and in the corresponding R code.

The IRS migration data are an incredible tool for undestanding migration. By providing these data in a readily available format, we hope to facilitate their use in descriptive, exploratory, and analytical analyses of migration in the United States through the use of administrative data. This data is particularly useful for understanding migration as a spatial entity and for investigating the evolution of migration systems over time.

>>>>

---
output: 
  pdf_document:
    number_sections: true
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: LATEX/svm-latex-ms.tex
title: "IRS County-to-County Migration Data"
thanks: "The data and code that supports the creation of this data are available in the Supplementary Materials and online at  https://github.com/mathewhauer/IRS-migration-data."
author:
  - name: Mathew E. Hauer ^1,2^*
    affiliation: Florida State University
  - name: James M. Byars ^3^
    affiliation: University of Georgia
# abstract: ""
keywords: "Migration, IRS, Data, R"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
#fontfamily: mathpazo
fontsize: 12pt
spacing: single
bibliography:  LATEX/mybibfile.bib
biblio-style: apsr
# use apsr or nature
header-includes:
- \usepackage[all]{nowidow}
- \usepackage{rotating}
- \usepackage{fancyhdr}
- \usepackage{multirow}
- \usepackage{tabularx}
- \pagestyle{fancy}
- \fancyhead{}
- \fancyhead[CE,CO]{IRS Migration Data}

@@

**OBJECTIVE**: To produce a single, flat data file containing complete county-to-county IRS migration flow data and to make the computer code used to process the migration data available.

@@

**CONTRIBUTION**: To encourage and facilitate the use of this data, we provide a single, standardized, flat data file containing county-to-county migration flows for the period 1990-2010 and provide the full R script to download, process, and flatten the IRS migration data.

@@

# Introduction
Migration flow data (ie, the number of migrants from location *i* to location *j*) are typically difficult to obtain information  despite their importance [@willekens2016international; @rogers2010indirect]. Migration scholars typically focus on cross-border, national migration flow data and recent country-to-country migration data are vital for understanding migration processes [@abel2014quantifying;@abel2017estimates;@abel2013estimating]. However, there is growing demonstrated importance surrounding subnational migration flows [@sorichetta2016mapping; @curtis2015recovery].

In the United States, subnational migration flow is available from three primary sources, depending on time period: the Decennial Census, the American Community Survey, and the Internal Revenue Service's (IRS) county-to-county migration data. The IRS migration data are a pioneering use of administrative records to estimate demographic processes and are available on an annual basis since 1990. Because of the annual availability, relatively long time series, large universe due to the administrative records, and long history of use, the IRS data are an attractive data source for conducting migration research in the United States (e.g. [@curtis2015recovery;@molloy2011internal;@shumway2001spatial;@frey2009great]). Unfortunately, these data exist in seven legacy formats, split between 2,000+ data files making analysis with this data rather burdensome and has likely hindered the widespread adoption of this valuable resource for US migration scholarship.

To encourage and facilitate the use of this tremendous migration resource, we make two contributions: (1) we publish a single, flat, standardized data file containing all county-to-county migration flows for the period 1990-2010, and (2) we publish the open-source R code used to process the IRS data into the single, flat, standardized data file for reproducibility, transparency, and educational purposes. By publishing both the migration data itself and the R code to collate, process, and flatten the IRS migration data, our hope is to save time and effort for other migration scholars and facilitate the use of this data. Scholars who wish to use these data should still familiarize themselves with the strengths and weaknesses, idiosyncrasies, and design of these data (see [@gross2005internal; @engels1981measuring;@franklin2006pandora;@pierce2015soi] for discussions on the IRS data) and with the procedures outlined in this document and in the corresponding R code^[The R code used to produce these data is available in the **Supplementary Materials** and can also be found in an online repository located at https://github.com/mathewhauer/IRS-migration-data].

We have attempted to introduce as little post-processing as possible to process the data into a common format. US Counties are fairly stable geographic units but some changes in county boundaries, names, and FIPS codes due occasionally occur^[The Federal Information Processing Standard Publication (FIPS) is a 5-digit code used to uniquely identify US counties and county equivalents.]. To try and keep as close to the original data fidelity as possible, we did not recode any geographic changes and present the IRS migration data as-is. For instance, Broomfield County, Colorado (FIPS 07014) was created out of parts of Adams, Boulder, Jefferson, and Weld counties in 2001 and thus has data only after 2002. Uses should be aware of any changes in county boundaries, county names, or FIPS changes that could substantially alter any analysis of this data^[More detailed information about county boundary, name, or FIPS changes can be found at the following locations https://www.census.gov/geo/reference/county-changes.html
http://www.nber.org/asg/ASG_release/County_City/FIPS/FIPS_Changes.pdf
https://www.cdc.gov/nchs/data/nvss/bridged_race/County_Geography_Changes.pdf
https://www.ddorn.net/data/FIPS_County_Code_Changes.pdf].

The following document is organized as follows. First, we describe the IRS county-to-county migration data to provide an overview of the data for scholars who might be unfamiliar with the IRS migration data. Second, we provide usage notes that provide important information that may assist other researchers who use our data. Third, we describe our single, flat, standardized file and document important nuances in the raw IRS migration data. Finally, we describe parts of the R code used to download the IRS migration data and process it into a common format.

The IRS migration data are an incredible tool for understanding migration. By providing these data in a readily available format and the subsequent open-source computer code used to process these data, we hope to facilitate their use in descriptive, exploratory, and analytical analyses of migration in the United States through the use of administrative data. This is data is particularly useful for understanding migration as a spatial entity and for investigating the evolution of migration systems over time.

# IRS Migration Data
The IRS began using tax data to estimate migration in the 1970s and 1980s [@engels1981measuring;@franklin2006pandora] and began releasing migration data in 1990. The IRS uses individual federal tax returns, matches these individual returns between two tax years (for instance tax year 2000 and tax year 2001), and identifies both migrants and non-migrants. Beginning with tax year 1991 (migration year 1990), the IRS produces these data in conjunction with the US Census Bureau using the IRS Individual Master File which contains every Form 1040, 1040A, and 1040EZ. Migration is identified when a current years' tax form contains an address that is different from the matched preceding years' return. A non-migrant is identified when there is no change in address between two years. For the 2002 tax year, the IRS migration data contained approximately 130 million returns [@gross2005internal].

The annual series of county-to-county migration data cover 95 to 98 percent of the tax filing universe (or approximately 87% of US households [@molloy2011internal]) and their dependents making these data the largest migration data source for count flows between counties in the United States. But since the IRS derives migration information from tax filings, those who do not file taxes are most likely to be underrepresented in the migration data [@gross2005internal; @dewaard2016population], namely undocumented populations, the poor, the elderly, and college students [@gross2005internal]. However, the overwhelming majority of householders file US tax returns in the United States [@molloy2011internal].

The IRS reports a number of important variables in their data. They identify both the origin and destination counties; the number of total migrants who moved from county $i$ to county $j$; the number of non-migrants in origin county $i$; and the numbers of tax returns or filers associated with those moves (roughly analogous to the number of households and listed as the `returns` field in the raw data) and the numbers of tax exemptions associated with those moves (roughly analogous to the number of individuals and listed as the `exemptions` field in the raw data).

Between 1990 and 2010, the IRS processed the county-to-county migration data using the same procedures. However, in 2011 the IRS introduced a new method for processing the migration data and introduced "enhancements" to improve the overall quality of the data [@pierce2015soi]. The IRS introduced three major changes. First, they began basing migration on a full year of data as opposed to a partial year of data. To meet Census Bureau deadlines, the IRS processed all income tax returns filed before the end of September and did not process the returns that were filed between the end of September and the end of the calendar year. Beginning with migration year 2011, the IRS included the approximately 4% of returns that are filed between the end of September and December 31, allowing the IRS to produce a full calendar years' worth of migration. Second, the IRS improved the year-to-year matching, increasing the number of matched returns by 5 percent. Third, the IRS began tabulating gross migration at the US State level by size of adjusted gross income (AGI) and the age of the primary taxpayer.

These changes to the processing of returns create a break in the historic time series. For this reason, we limit the data we process to the period 1990-2010, the last year before the new processing rules. If a scholar wishes to processes any IRS migration data after 2010, the R code that we provide can be easily adapted to do so.

# Usage Notes
The dataset generated here provides detailed county-to-county migration data based on administrative records. Users of these data should be aware that although the data have been prepared in a transparent manner with documentation of their creation and post-processing, and with open-source computer code, little was done to post-process the data to correct any possible inconsistencies or errors. These data should be used only with full awareness of the inherent limitations of the IRS migration data and with the knowledge of the procedures outlined in this document and in the corresponding R code. Caveat emptor -- users beware.

Users should be aware of several limitations of the IRS data. Namely, that any origin-destination pair with fewer than 10 tax filers is censored or suppressed by the IRS for privacy reasons. We have collected these censored flows into a unique FIPS code (FIPS 99999) by subtracting all uncensored flows from the total number of migrants. Any origin-destination pair with fewer than 10 tax filers over the entire period is thus excluded from the final datafile since no data would be recorded in the IRS datafile due to censoring.

Users should also be mindful of possible geographic changes to county boundaries that could impact the data.

The county migration data we present come from the `exemptions` field of the IRS migration data. The original IRS migration data contains two consistent fields across all years of data: a `returns` field and an `exemptions` field. Returns are the number of tax returns filed while exemptions are a proxy for the members of the household. This was done to better mimic the number of individuals migrating rather than the number of households.

**Table 1** demonstrates the general structure of our flat migration data file.

# Data Processing
The IRS migration data for the period 1990-2010 are available in seven legacy formats. **Table 2** summarizes some of the similarities and differences in these formats. For every year, the IRS publishes approximately 104 data files. (52 state entities by in/out-migration. Some years contain .csv and .dat summary files.) The underlying file organization, file format, naming schema, and coding can differ between these legacy formats. Migration years 1990 and 1991 are available as fixed-width text files, while 1992-2010 are available as excel files. For years 1990-2003, the IRS separated in/out migration into separate folders while 2004-2010 are published in a single folder. Each legacy format utilizes a different file naming scheme as well, making pattern matching of file names (called grepping) difficult. Importantly, the IRS treats non-migrants and total migrants differently in the seven legacy formats. For 1990 and 1991, the IRS simply has a field that reads "County Non-Migrants" for non-migrants; for 1992-1994, the IRS introduced a State code 63 but two difference County codes (010 for 1992 and 1994 and 050 for 1993) creating a 5-digits FIPS code of 63010 or 63050. After 1995, the IRS smartly set the origin FIPS equal to the destination FIPS for non-migrants. Lastly, Total Migrants are treated differently too. For 1990 and 1991, the destination field simply reads "Total Migration." For 1992-1994, the IRS introduced a State Code 00 and county code 001 for total migrants. After 1995, the IRS used State Code 96 and county Code 000 for a combined 5-digit FIPS code of 96000.

The differences described above and in **Table 2** are only some of the differences that are of interest to the data we produce here. Total Migrants, ie FIPS 96000 for migration data after 1995, is also broken down into Total Mig - US (FIPS 97000), Total Mig - US Same State (FIPS 97001), Total Mig - US Diff St (FIPS 97003), and Total Mig - Foreign (FIPS 98000). The IRS did not code these migration flows in this manner for all years, and in some cases (such as Total Mig - Foreign) migration flows are not reported. For simplicity and data continuity purposes, we simply create a new origin/destination (FIPS 99999) that contains all unspecified migration flows. We do this by subtracting the number of enumerated migrants (the migration flows with greater than 10 migrants) from the total number of migrants. This way, the sum of all enumerated migrants in our dataset will equal the total number of migrants in the IRS dataset.

@@

\begin{table}
\caption{Select differences in the file formats, file organizations, naming, and treatment of various migration statistics.}
\label{my-label}
\resizebox{\textwidth}{!}{%
\begin{tabular}{llllll}
\hline
\textbf{Years} & \textbf{Data Format} & \textbf{File Organization} & \textbf{Sample File naming} & \textbf{\begin{tabular}[c]{@{}l@{}}Coding of \\ non-migrants\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Coding of \\ Total Migrants\end{tabular}} \\ \hline
\multicolumn{1}{l|}{1990-1991} & \multicolumn{1}{l|}{txt} & \multicolumn{1}{l|}{\multirow{4}{*}{Separate in/out migration}} & \multicolumn{1}{l|}{C9091alo.txt} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Destination field reads\\  `County Non-Migrants'\end{tabular}} & \begin{tabular}[c]{@{}l@{}}Destination field reads \\ `Total Migration'\end{tabular} \\ \cline{2-2} \cline{4-6} 
\multicolumn{1}{l|}{1992, 1994} & \multicolumn{1}{l|}{\multirow{6}{*}{xls}} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{C9293Alo.xls} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}State code = 63,\\ County code = 010\end{tabular}} & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}State code = 00,\\ County code = 001\end{tabular}} \\ \cline{4-5}
\multicolumn{1}{l|}{1993} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{co934alo.xls} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}State code = 63,\\ County code = 050\end{tabular}} &  \\ \cline{4-6} 
\multicolumn{1}{l|}{1995-2003} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{co956alor.xls} & \multicolumn{1}{l|}{\multirow{4}{*}{\begin{tabular}[c]{@{}l@{}}Origin FIPS = \\ Destination FIPS\end{tabular}}} & \multirow{4}{*}{\begin{tabular}[c]{@{}l@{}}State code = 96,\\ County = 000\end{tabular}} \\ \cline{3-4}
\multicolumn{1}{l|}{2004-2006} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{\multirow{3}{*}{Single folder}} & \multicolumn{1}{l|}{co0405ALo.xls} & \multicolumn{1}{l|}{} &  \\ \cline{4-4}
\multicolumn{1}{l|}{2007-2008} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{co0708oAl.xls} & \multicolumn{1}{l|}{} &  \\ \cline{4-4}
\multicolumn{1}{l|}{2009-2010} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{co0910oAL.xls} & \multicolumn{1}{l|}{} &  \\ \hline
\end{tabular}%
}
\end{table}

## R Code

@@

## Data Download
The script `001-download_data.R` will download and unzip the migration data from the IRS' websites into a folder standardized format into subdirectory `MigData/.` The IRS data is in two primary formats: 1990-2003 and 2004-onward. There are eight files that the IRS includes in their zip archives that contain no data (these are in years 1998, 1999, 2000, and 2001). After being downloaded and unzipped, these files are then deleted. If they are not deleted, they will cause the subsequent `for loops` to fail in the next section. These files do not contain any migration information, their names suggest they represent aggregation of migration flows (for example 'co990usi.xls' suggests county 99-2000 US in-migration), and we are unsure exactly why these files are included or their purpose.

## Data Processing
The third and final section contains several `foreach` parallel processing loops to process the seven legacy formats into a common data format. These files are then row-bound using `rbindlist` and transformed into a 'short' data frame. **Table 1** demonstrates the general file layout. In- and out-migration files are processed separately and only unique dyadic flows are kept in the file flat file.
